---
title: "Foundations Project 2"
author: "Hunter Ratliff, @HunterRatliff1"
date: "10/16/2019"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, message=F, warning=F)
require(tidyverse)
require(ggthemes)
require(scales)
require(lubridate)
require(googlesheets)
```

```{r read_data}

# Read in data from table 17
mortality <- readr::read_csv("~/Google Drive/'UTMB/- MPH/[6335] Foundations of Public Health/Project 2 [Foundations PH, 6335]/Table17_Cancer.csv") %>% select(-X1) %>%
  mutate(year = as.numeric(year))
# mortality <- googlesheets::gs_title("table017") %>%
#   gs_read_csv() %>%
#   gather("year", "deaths", `1950`:`2016`) %>%
#   filter(Disease %in% c("Colon, rectum, and anus", "Malignant neoplasms", "All causes"))
  

# read in Healthy People data
HP <- readxl::read_excel("~/Google Drive/'UTMB/- MPH/[6335] Foundations of Public Health/Project 2 [Foundations PH, 6335]/C-16-data.xlsx") %>%
  mutate(PopTitle = str_replace(PopTitle, "Hispanic or Latino", "Latino")) %>%
  mutate(PopTitle = str_replace(PopTitle, " or African American", "")) %>%
  mutate(PopTitle = str_replace(PopTitle, "American Indian or Alaska Native", "AmIAK")) %>%
  
  # filter some vars
  filter(PopTitle!="White only") %>%
  filter(PopTitle!="Not Latino") %>%
  filter(PopTitle!="Black only, not Latino") %>%
  filter(PopTitle!="50-64 years") 


SEER <- readr::read_csv("~/Google Drive/'UTMB/- MPH/[6335] Foundations of Public Health/Project 2 [Foundations PH, 6335]/SEER.csv", skip = 3) %>%
  select(Sex:`AA Rate Upper Confidence Interval`) %>%
  filter(`Rate Type`=="Observed Rates") %>%
  select(-`Rate Type`) %>%
  rename(Race=`Race/Ethnicity`) %>%
  mutate(
    Race = str_replace(Race, "American Indian / Alaska Native", "AmIAN"),
    Race = str_replace(Race, "Asian / Pacific Islander", "Asian"),
    Race = str_replace(Race, "includes", "inc")
  ) %>%
  mutate(
    `Age-adjusted (AA) Rate per 100,000` = na_if(`Age-adjusted (AA) Rate per 100,000`, "-"),
    `AA Rate Lower Confidence Interval`  = na_if(`AA Rate Lower Confidence Interval`, "-"),
    `AA Rate Upper Confidence Interval`  = na_if(`AA Rate Upper Confidence Interval`, "-")
  ) %>%
  mutate(
    `Age-adjusted (AA) Rate per 100,000` = as.numeric(`Age-adjusted (AA) Rate per 100,000`),
    `AA Rate Lower Confidence Interval` = as.numeric(`AA Rate Lower Confidence Interval`),
    `AA Rate Upper Confidence Interval` = as.numeric(`AA Rate Upper Confidence Interval`)
  )
names(SEER) <- c("Sex", "Race", "Age", "Year", "Rate", "lower_ci", "upper_ci")

# Select mutually exclusive age categories
SEER <- SEER %>%
  mutate(Age = str_trim(str_replace(Age, "Ages", ""))) %>%
  filter(Age %in% c("< 50", "50-64", "65-74", "75+", "All")) %>%
  mutate(Race = str_trim(str_replace(Race, "\\(inc Hispanic\\)", ""))) 


SEER_mortality <- readr::read_csv("~/Google Drive/'UTMB/- MPH/[6335] Foundations of Public Health/Project 2 [Foundations PH, 6335]/SEER_Mortality.csv", skip = 3) %>%
  rename(Race=`Race/Ethnicity`) %>%
  mutate(
    Race = str_replace(Race, "American Indian / Alaska Native", "AmIAN"),
    Race = str_replace(Race, "Asian / Pacific Islander", "Asian"),
    Race = str_replace(Race, "includes", "inc")
  ) %>%
  mutate(
    `Age-adjusted (AA) Rate per 100,000` = na_if(`Age-adjusted (AA) Rate per 100,000`, "-"),
    `AA Rate Lower Confidence Interval`  = na_if(`AA Rate Lower Confidence Interval`, "-"),
    `AA Rate Upper Confidence Interval`  = na_if(`AA Rate Upper Confidence Interval`, "-")
  ) %>%
  mutate(
    `Age-adjusted (AA) Rate per 100,000` = as.numeric(`Age-adjusted (AA) Rate per 100,000`),
    `AA Rate Lower Confidence Interval` = as.numeric(`AA Rate Lower Confidence Interval`),
    `AA Rate Upper Confidence Interval` = as.numeric(`AA Rate Upper Confidence Interval`)
  )
names(SEER_mortality) <-  c("Sex", "Race", "Age", "Year", "Rate", "lower_ci", "upper_ci")
# Select mutually exclusive age categories
SEER_mortality <- SEER_mortality %>%
  mutate(Age = str_trim(str_replace(Age, "Ages", ""))) %>%
  filter(Age %in% c("< 50", "50-64", "65-74", "75+", "All")) %>%
  mutate(Race = str_trim(str_replace(Race, "\\(inc Hispanic\\)", ""))) 

```

# Objective

Briefly describe the burden of disease (magnitude, time trends, group differences) associated with the health outcome within the target population.

# Data sets

**HP:** Healthy People 2020 dataset from objective [C-16](https://www.healthypeople.gov/2020/data-search/Search-the-Data#objid=4054;). Outcome is percent of adults receiving colorectal cancer screening based on the most recent guidelines (age-adjusted, 50–75 years)

**Table 17**: Mortality data from earlier in the class

**SEER**: SEER data showing incidence rates of diagnosis of colon and rectum cancer, derived from [this table](https://seer.cancer.gov/explorer/application.php?site=20&data_type=1&graph_type=2&compareBy=sex&chk_sex_3=3&chk_sex_2=2&chk_race_1=1&chk_age_range_1=1&chk_data_type_1=1&advopt_precision=1&advopt_display=2&showDataFor=race_1_and_age_range_1_and_data_type_1)

# Magnitude & time trends

## By Mortality

```{r Mag-Mortality-df}
df <- mortality %>%
  filter(year>=1980) %>%
  mutate(Disease = recode(Disease,
                          "All causes"         = "Total",
                          "Malignant neoplasms"= "All Cancers",
                          "Colon, rectum, and anus"="CRC")) %>%
  spread(Disease, deaths) %>%
  mutate(
    `All Cancers` = CRC/`All Cancers`,
    Total         = CRC/Total
  )
```


```{r Mag-Mortality-1}
# df %>%
#   # Compare all people
#   filter(Category=="All persons") %>%
#   ggplot(aes(x=year, y=CRC)) +
#   geom_point(alpha=0.25) +
#   geom_line() +
#   theme_bw() +
#   labs(title="Age-adjusted rates of CRC mortality",
#        y="Deaths per 100k",
#        caption="Source: Table 17 National Vital Statistics System")
# 
# 
SEER_mortality %>%
  # Compare all people
  filter(Age=="All") %>%
  filter(Sex!="Both Sexes") %>%
  ggplot(aes(x=Year, y=Rate, group=Race)) +
  geom_ribbon(aes(ymin=lower_ci, ymax=upper_ci), alpha=0.5, color="grey") +
  # geom_point(alpha=0.25) +
  geom_line(aes(color=Race)) +
  theme_bw() +
  scale_color_manual(values=c(
    "White"="#268bd2",
    "Black"="#dc322f",
    "All Races"="#3C3C3C")) +
  guides(fill=F) +
  facet_wrap("Sex") +
  theme(legend.position = "bottom") +
  labs(title="Age-adjusted rates of CRC mortality",
       subtitle = "With 95% CIs",
       y="Deaths per 100k",
       caption="Source: SEER")
```

```{r survival5}
tibble(
  Stage = c("All Stages", "Localized", "Regional", "Distant"),
  Surv  = c(0.584, 0.869, 0.661, 0.11),
  lower_ci = c(0.576, 0.857, 0.644, 0.1),
  upper_ci = c(0.592, 0.88, 0.677, 0.12)
) %>%
  ggplot(aes(x=Stage, y=Surv, ymin=lower_ci, ymax=upper_ci)) +
  geom_col(aes(fill=Stage)) +
  geom_errorbar(width=0.5) +
  scale_x_discrete (limits = c("All Stages", "Localized", "Regional", "Distant")) +
  scale_y_continuous(name="% Surviving at 5 Years", labels = percent) +
  theme_bw() +
  guides(fill=F) +
  labs(title="5-Year Relative Survival Rates, By Stage at Diagnosis",
       x="Stage at Diagnosis",
       caption = "Source: SEER")
```



```{r Mag-Mortality-2}
df %>%
  # Compare all people
  filter(Category=="All persons") %>%
  ggplot(aes(x=year, y=Total)) +
  geom_point(alpha=0.25) +
  geom_line() +
  theme_bw() +
  scale_y_continuous(labels = percent) +
  labs(title="Proportion of all cause mortality",
       subtitle = "Proportionate mortality of CRC",
       y="CRC deaths / Total deaths\n(all cause)",
       caption="Source: Table 17 National Vital Statistics System")
```

```{r Mag-Mortality-3}
df %>%
  # Compare all people
  filter(Category=="All persons") %>%
  ggplot(aes(x=year, y=`All Cancers`)) +
  geom_point(alpha=0.25) +
  geom_line() +
  theme_bw() +
  scale_y_continuous(labels = percent) +
  labs(title="Proportion of all cancer deaths",
       subtitle = "Proportionate mortality of CRC",
       y="CRC deaths / Total cancer deaths",
       caption="Source: Table 17 National Vital Statistics System")
```


## By incidence

```{r Mag-SEER}
SEER %>%
  filter(Age!="All") %>%
  
  filter(Race=="All Races") %>%
  filter(Sex=="Both Sexes") %>% 
  ggplot(aes(x=Year, y=Rate, ymax=upper_ci, ymin=lower_ci)) +
  geom_line(aes(color=Age)) +
  theme_bw() +
  scale_color_manual(values=c(
    "< 50"="#4D9221", "50-64"= "#A1D76A",
    "65-74"= "#E9A3C9", "75+"="#C51B7D")) +
  
  labs(title="Rates of CRC diagnosis, by age",
       subtitle="Age-adjusted incidence",
       y="Incidence of CRC per 100k\n(age-adjusted)",
       caption = "Source: SEER")
  
```



# Group differences

## By mortality

```{r Group-Mortality-1}
mortality %>%
  filter(Disease=="Colon, rectum, and anus") %>%
  filter(year==2016) %>%
  filter(Category!="Black, non-Hispanic") %>%
  filter(Category!="White") %>%
  filter(CategoryType!="Race") %>%
  # View()
  ggplot(aes(x=Category, y=deaths, fill=Category)) +
  geom_col() +
  theme_bw() +
  coord_flip() +
  
  scale_fill_manual(values=c("All persons"="grey", "Males"="#0084D1", "Females"="#FFD320")) +
  guides(fill=F) +
  # facet_wrap("CategoryType", scale="free_y") +
  labs(title="CRC mortality, by Sex",
       subtitle="Age-adjusted, 2016",
       y="Deaths per 100k", x="",
       caption="Source: Table 17 National Vital Statistics System")
```

```{r Group-Mortality-2}
mortality %>%
  filter(Disease=="Colon, rectum, and anus") %>%
  filter(year==2016) %>%
  filter(Category!="Black, non-Hispanic") %>%
  filter(Category!="White") %>%
  filter(CategoryType!="Sex") %>%
  # View()
  ggplot(aes(x=Category, y=deaths, fill=Category)) +
  geom_col() +
  theme_bw() +
  coord_flip() +
  scale_fill_manual(values=c(
    "White, non-Hispanic"="#268bd2",
    "Hispanic"="#d33682",
    "Black"="#dc322f",
    "Asian"="#859900",
    "AmIAN"="#b58900",
    "All persons"="#3C3C3C")) +
  guides(fill=F) +
  # facet_wrap("CategoryType", scale="free_y") +
  labs(title="CRC mortality, by race",
       subtitle="Age-adjusted, 2016",
       y="Deaths per 100k", x="",
       caption="Source: Table 17 National Vital Statistics System")
```

```{r Group-Mortality-Both}
mortality %>%
  filter(Disease=="Colon, rectum, and anus") %>%
  filter(year==2016) %>%
  filter(Category!="Black, non-Hispanic") %>%
  filter(Category!="White") %>%
  # filter(CategoryType!="All") %>%
  mutate(CategoryType = recode(CategoryType, "All"="Sex")) %>%
  mutate(Category = str_replace(Category, ", ", ",\n")) %>%
  
  ggplot(aes(x=Category, y=deaths, fill=Category)) +
  geom_col(color="black") +
  theme_bw() +
  coord_flip() +
  scale_fill_manual(values=c(
    "White,\nnon-Hispanic"="#268bd2",
    "Hispanic"="#d33682",
    "Black"="#dc322f",
    "Asian"="#859900",
    "AmIAN"="#b58900",
    "All persons"="grey",
    "Males"="#7cb5ec", "Females"="#e4d354")) +
  guides(fill=F) +
  facet_wrap("CategoryType", scale="free_y") +
  labs(title="CRC mortality, by Race & Sex",
       subtitle="Age-adjusted, 2016",
       y="Deaths per 100k", x="",
       caption="Source: Table 17 National Vital Statistics System")
```



## By incidence

```{r Group-SEER-1}
SEER %>%
  mutate(Race = str_replace(Race, "\\(inc Hispanic\\)", "")) %>%
  filter(Age=="All") %>%
  filter(Sex!="Both Sexes") %>% 
  filter(Race!="All Races") %>%
  filter(Race!="White") %>%
  # count(Race) %>% .$Race
  ggplot(aes(x=Year, y=Rate, ymax=upper_ci, ymin=lower_ci)) +
  # geom_ribbon(aes(group=Race), color="grey", alpha=0.25) +
  geom_line(aes(color=Race)) +
  theme_bw() +
  scale_color_manual(values=c(
    "Non-Hispanic White"="#D95F02",
    "Hispanic (any race)"="#1B9E77",
    "Black"="#E7298A",
    "Asian"="#7570B3",
    "AmIAN"="#66A61E",
    "Total"="#3C3C3C")) +
  guides(fill=F) +
  facet_wrap("Sex", ncol=1) +
  labs(title="Rates of CRC diagnosis, by group",
       subtitle="Age-adjusted incidence",
       y="Incidence of CRC per 100k\n(age-adjusted)",
       caption = "Source: SEER")
  
```

```{r Group-SEER-2}
SEER %>%
  mutate(Race = str_replace(Race, "\\(inc Hispanic\\)", "")) %>%
  mutate(Race = str_replace(Race, "Hispanic ", "Hispanic\n")) %>%
  filter(Year==2016) %>%
  filter(Age=="All") %>%
  # filter(Sex!="Both Sexes") %>% 
  filter(Race!="All Races") %>%
  filter(Race!="White") %>%
  ggplot(aes(x=Race, y=Rate, ymax=upper_ci, ymin=lower_ci)) +
  # geom_ribbon(aes(group=Race), color="grey", alpha=0.25) +
  geom_col(aes(fill=Sex), position = "dodge") +
  geom_errorbar(aes(group=Sex),  position = "dodge", alpha=0.5) +
  # facet_wrap("Sex") + 
  
  theme_bw() +
  theme(legend.position = "bottom") +
  scale_fill_manual(values=c("Both Sexes"="grey", "Male"="#FF2700", "Female"="#008FD5")) +
  # scale_fill_manual(values=c(
  #   "Non-Hispanic\nWhite"="#D95F02",
  #   "Hispanic\n(any race)"="#1B9E77",
  #   "Black"="#E7298A",
  #   "Asian"="#7570B3",
  #   "AmIAN"="#66A61E",
  #   "Total"="#3C3C3C")) +
  # guides(fill=F) +
  
  coord_flip() + #facet_wrap("Sex", scales="free_y") +
  labs(title="Rates of CRC diagnosis, by group",
       subtitle="Age-adjusted incidence, 2016",
       y="Incidence of CRC per 100k\n(age-adjusted)",
       caption = "Source: SEER")
```


# Screening rates

## Interesting groups:

By % of poverty line

```{r HP-Poverty}
HP %>%
  filter(Year==max(Year)) %>%
  filter(Grouping=="Percent_PovertyLine"|Grouping=="Total") %>%
  ggplot(aes(x=PopTitle, y=Estimate, ymax=UpperCI, ymin=LowerCI)) +
  geom_col(aes(fill=PopTitle)) +
  geom_errorbar(width=.5) +
  theme_bw() +
  scale_fill_manual(values=c(
    "Total"="grey",    "<100"="#99000D",
    "100-199"="#CB181D",  "200-399"="#EF3B2C",
    "400-599"="#FB6A4A", "600+"="#FC9272")) +
  guides(fill=F) +
  coord_flip() +
  labs(title="% who received CRC screening, by % poverty line",
       subtitle = "Objective C-16 (2015 data)",
       caption="Healthy People 2020, Objective C-16",
       y="Percent of group (age adjusted)", x="Income (as % of poverty line)")
```

By Race

```{r HP-Race}
HP %>%
  filter(Year==max(Year)) %>%
  filter(Grouping=="Race"|Grouping=="Total") %>%
  filter(PopTitle!="AmIAK only") %>%
  filter(PopTitle!="2 or more races") %>%
  mutate(PopTitle = str_replace(PopTitle, ", ", ",\n")) %>%
  # count(PopTitle) %>% .$PopTitle
  ggplot(aes(x=PopTitle, y=Estimate, ymax=UpperCI, ymin=LowerCI)) +
  geom_col(aes(fill=PopTitle)) +
  geom_errorbar(width=.5) +
  theme_bw() +
  scale_fill_manual(values=c(
    "White only,\nnot Latino"="#268bd2",
    "Latino"="#d33682",
    "Black only"="#dc322f",
    "Asian only"="#859900",
    "Total"="grey")) +
  guides(fill=F) +
  scale_x_discrete (limits = c("Asian only","Latino", "Black only",
                               "White only,\nnot Latino","Total")) +
  coord_flip() +
  labs(title="% who received CRC screening, by race",
       subtitle = "Objective C-16 (2015 data)",
       caption="Healthy People 2020, Objective C-16",
       y="Percent of group (age adjusted)", x="")
```

By Age

```{r HP-Age}
HP %>%
  filter(Year==max(Year)) %>%
  filter(Grouping=="Age"|Grouping=="Total") %>%
  ggplot(aes(x=PopTitle, y=Estimate, ymax=UpperCI, ymin=LowerCI)) +
  geom_col(aes(fill=PopTitle)) +
  geom_errorbar(width=.5) +
  theme_bw() +
  scale_fill_manual(values=c(
    "50-54 years"="#6BAED6", "55-64 years"= "#4292C6",
    "65-75 years"= "#2171B5",
    "Total"="grey")) +
  guides(fill=F) +
  coord_flip() +
  labs(title="% who received CRC screening, by age group",
       subtitle = "Objective C-16 (2015 data)",
       caption="Healthy People 2020, Objective C-16",
       y="Percent of group (age adjusted)", x="")
```

By insurance status

```{r HP-Insurance}
HP %>%
  filter(Year==max(Year)) %>%
  filter(Grouping=="Insurance"|Grouping=="Total") %>%
  ggplot(aes(x=PopTitle, y=Estimate, ymax=UpperCI, ymin=LowerCI)) +
  geom_col(aes(fill=PopTitle)) +
  geom_errorbar(width=.5) +
  theme_bw() +
  theme(legend.position = "bottom") +
  scale_fill_manual(name="",values=c(
    "Insured"="#4DAF4A",  "Private"="#984EA3",
    "Public"="#377EB8",   "Uninsured"="#E41A1C",
    "Total"="grey")) +
  coord_flip() +
  scale_x_discrete (limits = c("Private","Public",
                               "Uninsured", "Insured", 
                               "Total")) +
  labs(title="% who received CRC screening, by Insurance Status",
       subtitle = "Objective C-16 (2015 data)",
       caption="Healthy People 2020, Objective C-16",
       y="Percent of group (age adjusted)", x="")
```

By place of birth

```{r HP-Birthplace}
HP %>%
  filter(Year==max(Year)) %>%
  filter(Grouping=="Birthplace"|Grouping=="Total") %>%
  ggplot(aes(x=PopTitle, y=Estimate, ymax=UpperCI, ymin=LowerCI)) +
  geom_col(aes(fill=PopTitle)) +
  geom_errorbar(width=.5) +
  theme_bw() +
  scale_fill_manual(values=c(
    "Outside US"="#FFD320", "US"="#0084D1", "Total"="grey")) +
  guides(fill=F) +
  # coord_flip() +
  scale_x_discrete (limits = c("Outside US", "US",  "Total")) +
  labs(title="% who received CRC screening, by Birthplace",
       subtitle = "Objective C-16 (2015 data)",
       caption="Healthy People 2020, Objective C-16",
       y="Percent of group (age adjusted)", x="")
```

By education level

```{r HP-Edu}
HP %>%
  filter(Year==max(Year)) %>%
  filter(Grouping=="Edu"|Grouping=="Total") %>%
  mutate(PopTitle = str_replace(PopTitle, " degree", "\ndegree")) %>%
  ggplot(aes(x=PopTitle, y=Estimate, ymax=UpperCI, ymin=LowerCI)) +
  geom_col(aes(fill=PopTitle)) +
  geom_errorbar(width=.5) +
  theme_bw() +
  scale_fill_manual(values=c(
    "Advanced\ndegree"="#7A0177",   "4-year college\ndegree"="#AE017E",
    "Associates\ndegree"="#DD3497", "Some college"="#F768A1",
    "High school"="#FA9FB5",       "< High school"="#FCC5C0",
    "Total"="grey")) +
  guides(fill=F) +
  coord_flip() +
  scale_x_discrete (limits = c("Advanced\ndegree", "4-year college\ndegree",
                               "Associates\ndegree", "Some college",
                               "High school", "< High school",
                               "Total")) +
  labs(title="% who received CRC screening, by education",
       subtitle = "Objective C-16 (2015 data)",
       caption="Healthy People 2020, Objective C-16",
       y="Percent of group (age adjusted)", x="Education Level")

# RColorBrewer::brewer.pal(8, "RdPu") %>% show_col()
# RColorBrewer::brewer.pal(8, "RdPu")
```


# Survival plots

```{r}
Surv <- readr::read_csv("SEER_Surv.csv") %>%
  filter(SurvivalRate!="-") %>%
  mutate(
    Sex = recode(Sex, "Both Sexes"="Both"),
    Time = str_trim(str_replace(Time, "years", "year")),
    Time = str_trim(str_replace(Time, "year", "")),
  ) %>%
  
  mutate(
    Lower_CI = ifelse(Time=="Diagnosis", 100, Lower_CI),
    Upper_CI = ifelse(Time=="Diagnosis", 100, Upper_CI),
    Time     = ifelse(Time=="Diagnosis", "0", Time)
  ) %>%
  
  
  # Handle ages
  mutate(Age = str_trim(str_replace(Age, "Ages", ""))) %>%
  filter(Age %in% c("< 50", "50-64", "65-74", "75+", "All")) %>%
  
  # handle race
  mutate(
    Race = str_trim(str_replace(Race, " \\(includes Hispanic\\)", "")),
    Race = str_trim(str_replace(Race, " \\(any race\\)", "")),
    Race = recode(Race, "Asian / Pacific Islander"="Asian"),
    Race = recode(Race, "American Indian / Alaska Native"="AmIAN"),
    Race = recode(Race, "All Races"="All")
  ) %>%
  filter(Race!="White") %>%
  
  # Make numeric
  mutate(
    Time         = as.numeric(Time),
    SurvivalRate = as.numeric(SurvivalRate),
    SE           = as.numeric(SE),
    Lower_CI     = as.numeric(Lower_CI),
    Upper_CI     = as.numeric(Upper_CI)
  )
```


```{r surv_plot}
Surv %>%
  filter(Sex=="Both", Race=="All", Age=="All", StageAtDx!="Unstaged") %>%
  ggplot(aes(x=Time, y=SurvivalRate, ymin=Lower_CI, ymax=Upper_CI, 
             group=StageAtDx)) +
  geom_ribbon(alpha=0.5, color="grey") +
  geom_line(aes(color=StageAtDx)) +
  geom_point(aes(color=StageAtDx)) +
  # facet_wrap("Age") +
  theme_bw() +
  theme(legend.position = "bottom") +
  scale_color_discrete(name="Stage at\ndiagnosis") +
  scale_y_continuous(breaks=seq(0,100,20),limits = c(0,100)) +
  labs(x="Years (since diagnosis)", y="Percent surviving",
       title="Survival by stage at diagnosis",
       caption = "Source: SEER")
```

```{r}
tibble(
  Race      = c("All", "AmIAN", "Asian", "Black", "Hispanic", "Non-Hispanic White"),
  Localized	= c(39.3,	36.4,	39.0,	38.0,	37.6,	39.5),
  Regional  = c(34.3,	34.8,	35.4,	31.1,	35.1,	34.8),
  Distant   = c(20.2,	22.8,	19.2,	24.1,	21.0,	19.7),
  Unstaged  = c(6.2, 	6.0, 	6.4, 	6.8, 	6.2, 	6.0)
) %>%
  gather("Stage", "Percent", -Race) %>%
  mutate(Percent = Percent/100,
    Stage   = factor(Stage, levels=c("Localized", "Regional", "Distant", "Unstaged"))) %>%
  filter(Race != "All", Race != "AmIAN", Stage != "Unstaged") %>%
  ggplot(aes(x=Stage, y=Percent, fill=Race)) +
  geom_col(position = "dodge", color="white") +
  theme_bw() +
  scale_fill_manual(values=c(
    "Non-Hispanic White"="#268bd2",
    "Hispanic"="#b58900",
    "Black"="#dc322f",
    "Asian"="#859900")) +
  scale_y_continuous(labels=percent) +
  labs(x="", y="% of cases (for each race)",
       title="Stage at diagnosis, by race",
       caption="Source: SEER")

```


