---
title: "Neighborhood Assessment"
author: "Hunter Ratliff"
date: "1/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F, warning=F, message=F)

library(ggthemes)
library(scales)
library(tidycensus)
library(tigris)
library(leaflet)
library(tidyverse)

options(tigris_use_cache = TRUE)

# Tracts only on the island
island_tracts <- c(48167724101, 48167724200, 48167724000, 48167724300, 48167724500, 48167724600,
                   48167724400, 48167724800, 48167724700, 48167724900, 48167725100, 48167725000,
                   48167725200, 48167725300, 48167725400, 48167725500, 48167725800, 48167725600,
                   48167725700, 48167725900) # 48167726000
# get shape files
galvy <- tigris::tracts("TX", "Galveston", cb = T)
NB_tract <- tigris::tracts("TX", "Galveston", cb = T) %>%
  geo_join(tibble(GEOID = c(48167724600)), by="GEOID", how="inner") 

# # Check your work
# galvy %>%
#   geo_join(tibble(GEOID = island_tracts), by="GEOID", how="inner") %>%
#   leaflet() %>% addTiles() %>%
#   addPolygons(popup = ~GEOID)
```

```{r, eval=F}
v17 <- load_variables(2017, "acs5", cache = TRUE)
s17 <- load_variables(2017, "acs5/subject", cache = TRUE)

v17 %>% filter(str_detect(name, "B17015")) %>% 
  mutate(label = str_replace(label, "Estimate!!", "")) %>%
  tidyr::separate(label, into=c("1","2", "3", "4"), 
                  extra="merge", fill="right", sep="!!") %>%
  # mutate(label = str_replace(label, "!!", "\n")) %>%
  View()
# B17015_002 Families below FPL in past 12 months
# B17015_025 Families above FPL in past 12 months

s17 %>% filter(str_detect(name, "S1101")) %>% View()

## Tables
# B24021: Median Earnings by Industry
# B23020: Mean hours worked
# B19058: SNAP


v17 %>% filter(str_detect(name, "B24021")) %>% View()

```

Data sources: 2017 5-year ACS

```{r poverty}
# v17 %>% filter(str_detect(name, "B17015")) %>% 
#   mutate(label = str_replace(label, "Estimate!!", "")) %>%
#   tidyr::separate(label, into=c("1","2", "3", "4"), 
#                   extra="merge", fill="right", sep="!!") %>%
#   View()
poverty <- tidycensus::get_acs("tract", variables = c("B17015_002","B17015_025"),
                    state='TX', county = "Galveston", #geometry = TRUE,
                    summary_var = "B17015_001") %>%
  
  mutate(variable = case_when(
    variable == "B17015_002" ~ "Families below FPL", # in past 12 months
    variable == "B17015_025" ~ "Families above FPL",
    T                        ~ variable
  )) %>%
  filter(variable=="Families below FPL") %>%
  
  mutate(
    value = estimate/summary_est,
    value_moe = moe_prop(estimate, summary_est, moe, summary_moe)
  ) %>%
  filter(GEOID %in% island_tracts) %>%
  mutate(OurTract = if_else(GEOID==48167724600, "Yes", "No")) %>%
  ungroup() %>%
  mutate(NAME = str_replace(NAME, ", Galveston County, Texas", ""))
```

```{r single}
# s17 %>% filter(str_detect(name, "S1101")) %>% 
#   mutate(label = str_replace(label, "Estimate!!", "")) %>%
#   tidyr::separate(label, into=c("1","2", "3", "4"), 
#                   extra="merge", fill="right", sep="!!") %>%
#   View()
single <- tidycensus::get_acs("tract", variables = c("S1101_C02_003","S1101_C03_003", "S1101_C04_003"),
                    summary_var = "S1101_C01_003", # total number of families
                    state='TX', county = "Galveston") %>%
  mutate(variable = case_when(
    variable == "S1101_C02_003" ~ "Married",
    variable == "S1101_C03_003" ~ "Single dad",
    variable == "S1101_C04_003" ~ "Single mom",
    T ~ variable
  )) %>%
  filter(variable!="Married") %>%
  group_by(GEOID, NAME, summary_est, summary_moe) %>%
  summarise(
    moe = round(moe_sum(moe, estimate = estimate)),
    estimate = sum(estimate)
  ) %>%
  mutate(
    value = estimate/summary_est,
    value_moe = moe_prop(estimate, summary_est, moe, summary_moe)
  ) %>%
  ungroup() %>%
  filter(GEOID %in% island_tracts) %>%
  mutate(OurTract = if_else(GEOID==48167724600, "Yes", "No")) %>%
  mutate(NAME = str_replace(NAME, ", Galveston County, Texas", "")) %>%
  mutate(variable = "Single-parent families") %>%
  select(GEOID, NAME, variable, 
         estimate, moe, summary_est, summary_moe,
         value, value_moe, OurTract)
```

```{r income}
# v17 %>% filter(str_detect(name, "B19013")) %>% View()   #median household income
income <- tidycensus::get_acs("tract", variables = c("B19013_001"),  # all races
                    state='TX', county = "Galveston") %>%
  mutate(variable = case_when(
    variable == "B19013_001" ~ "Median household income",
    T ~ variable
  )) %>%
  mutate(
    value = estimate,
    value_moe = moe
  ) %>%
  filter(GEOID %in% island_tracts) %>%
  mutate(OurTract = if_else(GEOID==48167724600, "Yes", "No")) %>%
  ungroup() %>%
  mutate(NAME = str_replace(NAME, ", Galveston County, Texas", ""))
```

```{r employment}
# v17 %>% filter(str_detect(name, "B23025")) %>% 
#   mutate(label = str_replace(label, "Estimate!!", "")) %>%
#   tidyr::separate(label, into=c("1","2", "3", "4"), 
#                   extra="merge", fill="right", sep="!!") %>%
#   View()   # Unemployed in civilian labor force
employment <- tidycensus::get_acs("tract", variables = c("B23025_005"), 
                                  summary_var = "B23025_003",  # total civilian labor force
                    state='TX', county = "Galveston") %>%
  mutate(variable = case_when(
    variable == "B23025_005" ~ "Unemployment",
    T ~ variable
  )) %>%
  mutate(
    value = estimate/summary_est,
    value_moe = moe_prop(estimate, summary_est, moe, summary_moe)
  ) %>%
  filter(GEOID %in% island_tracts) %>%
  mutate(OurTract = if_else(GEOID==48167724600, "Yes", "No")) %>%
  ungroup() %>%
  mutate(NAME = str_replace(NAME, ", Galveston County, Texas", ""))
```

```{r race}
# v17 %>% filter(str_detect(name, "B03002")) %>%
#   mutate(label = str_replace(label, "Estimate!!", "")) %>%
#   tidyr::separate(label, into=c("1","2", "3", "4"),
#                   extra="merge", fill="right", sep="!!") %>%
#   View()   # Race

race <- tidycensus::get_acs("tract", table="B03002", # whole table
                    summary_var = "B03002_001", 
                    state='TX', county = "Galveston") %>%
  mutate(variable = case_when(
    variable == "B03002_012" ~ "Hispanic",
    variable == "B03002_003" ~ "White",
    variable == "B03002_004" ~ "Black",
    variable == "B03002_006" ~ "Asian",
    variable %in% c("B03002_005", "B03002_007", "B03002_009") ~ "Other",
    T ~ "drop"
  )) %>%
  filter(variable!="drop") %>%
  group_by(GEOID, NAME, variable, summary_est, summary_moe) %>%
  summarise(
    moe = round(moe_sum(moe, estimate = estimate)),
    estimate = sum(estimate)
  ) %>%
  mutate(
    value = estimate/summary_est,
    value_moe = moe_prop(estimate, summary_est, moe, summary_moe)
  ) %>%
  ungroup() %>%
  filter(GEOID %in% island_tracts) %>%
  mutate(OurTract = if_else(GEOID==48167724600, "Yes", "No")) %>%
  mutate(NAME = str_replace(NAME, ", Galveston County, Texas", "")) %>%
  select(GEOID, NAME, variable, 
         estimate, moe, summary_est, summary_moe,
         value, value_moe, OurTract)
```




## Families living in poverty

Based on table [B17015](https://censusreporter.org/tables/B17015/), showing percent of familes who were below the federal poverty line (FPL) in the past 12 months

```{r poverty_summary}
overall <- poverty %>%
  summarise(
    moe = moe_sum(moe, estimate),
    summary_moe = moe_sum(summary_moe, summary_est),
    estimate = sum(estimate),
    summary_est = sum(summary_est)
  ) %>%
  mutate(
    value = estimate/summary_est,
    value_moe = moe_prop(estimate, summary_est, moe, summary_moe)
  ) %>%
  mutate(value = glue::glue("{round(value, 3)*100}% (+/- {round(value_moe, 3)*100}%)")) %>%
  .$value
  
ours <- poverty %>% filter(OurTract=="Yes") %>%
  mutate(value = glue::glue("{round(value, 3)*100}% (+/- {round(value_moe, 3)*100}%)")) %>%
  .$value

x <- list(Overall=overall, Ours=ours)
x
rm(overall, ours)
```

The pooled estimate of families living below the FPL in the past 12 months on the island is `r x[["Overall"]]` but our tract has `r x[["Ours"]]` families living below the FPL


```{r poverty_leaflet}
# factpal <- colorFactor(c("grey", "#FF2700"), df$OurTract)
pal <- colorNumeric(palette = "YlGnBu", domain = poverty$value)

geo_join(galvy, poverty, by="GEOID", how="inner") %>%
  leaflet() %>% addProviderTiles(providers$CartoDB.Positron) %>%
  addProviderTiles(providers$Stamen.TonerLabels) %>%
  addPolygons(data=NB_tract, color="red", fill=F, dashArray = "10", weight=4) %>%
  addPolygons(popup = ~as.character(round(100*value,1)), 
              color = ~pal(value),
              stroke=F,
              fillOpacity = 0.5) %>%
  addLegend("bottomright", pal = pal, values = ~value,
            title = "Families below FPL",
            labFormat = labelFormat(suffix="%",
              transform = function(x) 100 * x
            ),
            opacity = 1
  )
```

```{r poverty_plot}
poverty %>%
  ggplot(aes(x=reorder(NAME, value), y=value, fill=OurTract)) + 
  geom_col() + 
  geom_errorbar(aes(ymin=value-value_moe, ymax=value+value_moe), alpha=.5) +
  coord_flip() + scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("grey70", "#FF2700"), guide=F) + 
  labs(x="Census Tract", y="% of Families Below FPL in past year") +
  theme_bw()
```

```{r poverty_ggmap}
fortify(galvy, region="GEOID") %>%
  inner_join(poverty, by=c("id"="GEOID")) %>%
  mutate(OurTract=ifelse(OurTract=="Yes", "Yes", NA)) %>%
  ggplot(aes(long, lat, group=id, fill=value)) +
  geom_polygon() +
  geom_path(aes(color=OurTract)) +
  scale_fill_distiller(palette = "RdYlGn", direction = -1,
                       name="% families\nbelow FPL", labels=percent) +
  scale_color_manual(values = c("black"), guide=F) +
  coord_equal() + theme_map() +
  # theme(legend.position = "right")
  theme(legend.position = c(.71,.2))
```




## Single Parent Households

Based on table [S1101](http://factfinder2.census.gov/bkmk/table/1.0/en/ACS/16_1YR/S1101/0100000US), showing percent of familes who are single-parents

```{r single_summary}
overall <- single %>%
  summarise(
    moe = moe_sum(moe, estimate),
    summary_moe = moe_sum(summary_moe, summary_est),
    estimate = sum(estimate),
    summary_est = sum(summary_est)
  ) %>%
  mutate(
    value = estimate/summary_est,
    value_moe = moe_prop(estimate, summary_est, moe, summary_moe)
  ) %>%
  mutate(value = glue::glue("{round(value, 3)*100}% (+/- {round(value_moe, 3)*100}%)")) %>%
  .$value
  
ours <- single %>% filter(OurTract=="Yes") %>%
  mutate(value = glue::glue("{round(value, 3)*100}% (+/- {round(value_moe, 3)*100}%)")) %>%
  .$value

x <- list(Overall=overall, Ours=ours)
x
rm(overall, ours)
```


The pooled estimate of single-parent families on the island is `r x[["Overall"]]` but our tract has `r x[["Ours"]]` of families only having one parent in the household




```{r single_leaflet}
pal <- colorNumeric(palette = "YlOrRd", domain = single$value)

geo_join(galvy, single, by="GEOID", how="inner") %>%
  leaflet() %>% addProviderTiles(providers$CartoDB.Positron) %>%
  addProviderTiles(providers$Stamen.TonerLabels) %>%
  addPolygons(data=NB_tract, color="#FF2700", fill=F, dashArray = "10", weight=4) %>%
  addPolygons(popup = ~paste0(round(100*value,1),"% [MOE +/-", round(100*value_moe),"%]"), 
              color = ~pal(value),
              stroke=F,
              fillOpacity = 0.5) %>%
  addLegend("bottomright", pal = pal, values = ~value,
            title = "Single-parent families",
            labFormat = labelFormat(suffix="%",
              transform = function(x) 100 * x
            ),
            opacity = 1
  )
```

```{r single_plot}
single %>%
  ggplot(aes(x=reorder(NAME, value), y=value, fill=OurTract)) + 
  geom_col() + 
  geom_errorbar(aes(ymin=value-value_moe, ymax=value+value_moe), alpha=.5) +
  coord_flip() + scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("grey70", "#FF2700"), guide=F) + 
  labs(x="Census Tract", y="% Families with one parent") +
  theme_bw()
```

# Median household income

Based on table [B19013](https://censusreporter.org/tables/B19013/), showing median household income

```{r income_summary}
# overall <- income %>%
#   summarise(
#     moe = moe_sum(moe, estimate),
#     summary_moe = moe_sum(summary_moe, summary_est),
#     estimate = sum(estimate),
#     summary_est = sum(summary_est)
#   ) %>%
#   mutate(
#     value = estimate/summary_est,
#     value_moe = moe_prop(estimate, summary_est, moe, summary_moe)
#   ) %>%
#   mutate(value = glue::glue("{round(value, 3)*100}% (+/- {round(value_moe, 3)*100}%)")) %>%
#   .$value
#   
# ours <- income %>% filter(OurTract=="Yes") %>%
#   mutate(value = glue::glue("{round(estimate, 3)*100}% (+/- {round(moe, 3)*100}%)")) %>%
#   .$value
# 
# x <- list(Overall=overall, Ours=ours)
# x
# rm(overall, ours)
```



```{r income_leaflet}
pal <- colorNumeric(palette = "RdYlGn", domain = income$estimate)

geo_join(galvy, income, by="GEOID", how="inner") %>%
  leaflet() %>% addProviderTiles(providers$CartoDB.Positron) %>%
  addProviderTiles(providers$Stamen.TonerLabels) %>%
  addPolygons(data=NB_tract, color="black", fill=F, dashArray = "10", weight=2) %>%
  addPolygons(popup = ~paste0("$",estimate, "[MOE +/- $", moe, "]"), 
              color = ~pal(estimate),
              stroke=F,
              fillOpacity = 0.5) %>%
  addLegend("bottomright", pal = pal, values = ~estimate,
            title = "Median income",
            opacity = 1
  )

```




# Unemployment

Based on table [B23025](https://censusreporter.org/tables/B23025/), showing percent of civilian labor force who is unemployed

```{r employment_summary}
overall <- employment %>%
  summarise(
    moe = moe_sum(moe, estimate),
    summary_moe = moe_sum(summary_moe, summary_est),
    estimate = sum(estimate),
    summary_est = sum(summary_est)
  ) %>%
  mutate(
    value = estimate/summary_est,
    value_moe = moe_prop(estimate, summary_est, moe, summary_moe)
  ) %>%
  mutate(value = glue::glue("{round(value, 3)*100}% (+/- {round(value_moe, 3)*100}%)")) %>%
  .$value
  
ours <- employment %>% filter(OurTract=="Yes") %>%
  mutate(value = glue::glue("{round(value, 3)*100}% (+/- {round(value_moe, 3)*100}%)")) %>%
  .$value

x <- list(Overall=overall, Ours=ours)
x
rm(overall, ours)
```

The pooled estimate of unemployment on the island is `r x[["Overall"]]` but our tract has `r x[["Ours"]]` of the civilian labor force who is unemployed


# Race

Based on table [B03002](https://censusreporter.org/tables/B03002/), showing race & hispanic ethnicity

```{r}
df <- race %>%
  group_by(variable) %>%
  summarise(
    moe = moe_sum(moe, estimate),
    summary_moe = moe_sum(summary_moe, summary_est),
    estimate = sum(estimate),
    summary_est = sum(summary_est)
  ) %>%
  mutate(
    value = estimate/summary_est,
    value_moe = moe_prop(estimate, summary_est, moe, summary_moe)
  ) %>%
  select(variable, value, value_moe) %>%
  mutate(NAME = "Overall on island") %>%
  bind_rows(
    select(filter(race, OurTract=="Yes"), variable, value, value_moe, NAME)
  ) 
  # mutate(valueText = glue::glue("{round(value, 3)*100}% (+/- {round(value_moe, 3)*100}%)")) 

df %>%
  mutate(labelText = glue::glue("{round(value, 3)*100}%")) %>%
  mutate(labelText = ifelse(value>.04, labelText, NA)) %>%
  ggplot(aes(x=variable, y=value, group=NAME, label=labelText)) +
  geom_col(aes(fill=NAME), position = "dodge", color="black") +
  # geom_col(position = "dodge", fill="grey70", aes(alpha=ifelse(NAME=="Overall on island",1,0))) +
  # geom_errorbar(aes(ymin=value-value_moe, ymax=value+value_moe), alpha=.5) +
  coord_flip() + scale_y_continuous(labels = scales::percent) +
  geom_label(position = position_dodge(width = 1), hjust=1) +
  scale_fill_manual(values = c("#FF2700", "grey70"), name="Population") +
  labs(x="Race", y="% of total") +
  theme_bw()

df %>%
  mutate(value = glue::glue("{round(value, 3)*100}%")) %>%
  select(-value_moe) %>%
  spread(NAME, value) %>%
  arrange(desc(`Census Tract 7246`)) %>%
  knitr::kable()
  
```


